# Coursera - Data Science - Reproducible Research
## Assignment - 1  
***
### Loading and preprocessing the data  

The data is loaded using 'read.csv'.
```{r read}
x <- read.csv("activity.csv")
```

The date column was changed from factor class to date class.
```{r transform}
x$date <- as.Date(as.character(x$date))
```
***
### What is mean total number of steps taken per day?  

The subset of the initial dataframe which does not have NA's in the steps column.
```{r subset}
x_new <- x[!is.na(x$steps),]
```

Calcutale the total number of steps taken per day.
```{r average_x}
x_sum <- tapply(x_new$steps, x_new$date, sum)
```

Histogram(not barplot) plot of total number of steps taken per day.
```{r plot1}
hist(x_sum, xlab = "Days", main = "Total number of steps taken each day")
```

Caculating the mean and median of the total number of steps taken per day.
```{r x_mean_median}
x_mean <- mean(x_sum)
x_median <- median(x_sum)
```

Mean is `r x_mean` and median is `r x_median`.

***
### What is the average daily activity pattern?  
Average number of strps taken across all days in a given 5-minute interval. And a new variable 'intervals' which stores the intervals.
```{r intervalaverage}
x_avg <- tapply(x_new$steps, x_new$interval, mean)
intervals <- as.numeric(levels(factor(x_new$interval)))
```

Time series plot of the above:
```{r plot2}
plot(intervals, x_avg, type = "l", xlab = "Interval", ylab = "Average number of steps", main = "Average number of steps in each interval")
```

Max number of steps taken across all days in the 5-minutes interval.
```{r maxsteps}
max_step <- intervals[which(x_avg == max(x_avg))]
```
Max number of steps taken is `r max_step`.

***
### Imputing missing values  
Calculating total number of missing values.
```{r missingnumber}
missing <- sum(is.na(x$steps))
```
Total number of missing values are `r missing`.

I am using the mean of 5-minutes interval average to fill in the missing values in steps.

The new dataset 'y' is made.
```{r newdataset}
y <- x
y[which(is.na(x$steps)),]$steps <- x_avg
```

Time series plot of total number of steps taken each day of the new dataset.
```{r average_y}
y_sum <- tapply(y$steps, y$date, sum)
hist(y_sum, xlab = "Days", main = "Total number of steps taken each day")
```

Calculating mean and median of the new dataset.
```{r y_mean_median}
y_mean <- mean(y_sum)
y_median <- median(y_sum)
```

The new mean is `r y_mean` which is equal to the old mean `r x_mean`, therefore there is no change in the mean of the data.

The new median is `r y_median` which is greater than the old mean `r x_median`.

***
### Are there differences in activity patterns between weekdays and weekends?  
Creating a new variable for the dataset named 'day' which stores whether the day is 'weekday' or 'weekend'.
```{r day}
x_new$day <- ""
x_new[grep("S(at|un)", weekdays(x_new$date)),]$day <- "weekend"
x_new[!(x_new$day == "weekend"),]$day <- "weekday"
x_new$day <- factor(x_new$day, levels = c("weekend","weekday"))
```

Panel plot of average number of steps taken on all weekdays or weekend days.
```{r plot3}
par(mfrow = c(1,2))
x_avg_weekday <- tapply(x_new[x_new$day %in% "weekday",]$steps, x_new[x_new$day %in% "weekday",]$interval, mean)
x_avg_weekend <- tapply(x_new[x_new$day %in% "weekend",]$steps, x_new[x_new$day %in% "weekend",]$interval, mean)
plot(intervals, x_avg_weekday, type = "l", xlab = "Interval", ylab = "Average number of steps", main = "Weekdays", ylim = c(0,250))
plot(intervals, x_avg_weekend, type = "l", xlab = "Interval", ylab = "Average number of steps", main = "Weekends", ylim = c(0,250))
```
